---
title: "Draft 1"
output: html_notebook
---

# Loading Packages

```{r, warning = FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(usmap)
library(stargazer)
library(haven)
library(tigris)
library(rio)
library(standardize)
library(effsize)
library(ggthemes)
```

# Load and Clean Datasets

```{r, warning = FALSE, message = FALSE}
# Controls v2 from ACS
controls_2019 <- read_csv("controls_2019_v2.csv") %>%
  mutate(total_pop = SE_A00001_001) %>%
  mutate(percent_white = SE_A03001_002/total_pop) %>%
  mutate(percent_bachelors = SE_B12001_004/SE_B12001_001) %>%
  mutate(med_household_income = SE_A14006_001) %>%
  mutate(percent_hispanic = SE_B04001_010/total_pop) %>%
  select(Geo_FIPS, Geo_NAME, Geo_STATE, Geo_COUNTY, total_pop, 
         percent_white, percent_bachelors, med_household_income,
         percent_hispanic) %>%
  mutate(fips = as.numeric(Geo_FIPS))

cbp_2018 <- read_csv("cbp_2018.csv") %>%
  mutate(avg_wage = SE_T1200_001) %>%
  mutate(fips = as.numeric(Geo_FIPS)) %>%
  select(fips, avg_wage)

controls_2019 <- left_join(controls_2019, cbp_2018, by = "fips")
```

```{r, warning = FALSE, message = FALSE}
# Other Controls
poverty <- read_excel("PovertyEstimates.xls") %>%
  row_to_names(4) %>%
  clean_names() %>%
  mutate(fips = as.numeric(fip_stxt),
         poverty = as.numeric(pctpovall_2019)/100) %>%
  select(fips, poverty)

unemployment <- read_excel("Unemployment.xlsx") %>%
  row_to_names(4) %>%
  clean_names() %>%
  mutate(fips = as.numeric(fips_code),
         unemp = as.numeric(unemployment_rate_2019)/100) %>%
  select(fips, unemp)

controls_2019 <- controls_2019 %>%
  left_join(poverty, by = "fips") %>%
  left_join(unemployment, by = "fips")
```

```{r, warning = FALSE, message = FALSE}
# Import county migration data
c2c1519 <- import_list("c2c1519.xlsx")
c2c1216 <- import_list("c2c1216.xlsx")
c2c0812 <- import_list("c2c0812.xls")

# Rename columns
for (i in 1:52){
  colnames(c2c1519[[i]]) <- c("state_code_a", 
                                          "fips_a", 
                                          "state_code_b",
                                          "fips_b",
                                          "state_name_a",
                                          "county_name_a",
                                          "state_name_b",
                                          "county_name_b",
                                          "flow_ba",
                                          "flow_ba_moe",
                                          "flow_ab",
                                          "flow_ab_moe",
                                          "net_ba",
                                          "net_ba_moe",
                                          "gross_ab",
                                          "gross_ab_moe")
  colnames(c2c1216[[i]]) <- c("state_code_a", 
                                          "fips_a", 
                                          "state_code_b",
                                          "fips_b",
                                          "state_name_a",
                                          "county_name_a",
                                          "state_name_b",
                                          "county_name_b",
                                          "flow_ba",
                                          "flow_ba_moe",
                                          "flow_ab",
                                          "flow_ab_moe",
                                          "net_ba",
                                          "net_ba_moe",
                                          "gross_ab",
                                          "gross_ab_moe")
  colnames(c2c0812[[i]]) <- c("state_code_a", 
                                          "fips_a", 
                                          "state_code_b",
                                          "fips_b",
                                          "state_name_a",
                                          "county_name_a",
                                          "state_name_b",
                                          "county_name_b",
                                          "flow_ba",
                                          "flow_ba_moe",
                                          "flow_ab",
                                          "flow_ab_moe",
                                          "net_ba",
                                          "net_ba_moe",
                                          "gross_ab",
                                          "gross_ab_moe")
}

# Delete extraneous rows
for (i in 1:52){
  c2c1519[[i]] <- c2c1519[[i]] %>%
    filter(!row_number() %in% c(1, 2))
  c2c1216[[i]] <- c2c1216[[i]] %>%
    filter(!row_number() %in% c(1, 2))
  c2c0812[[i]] <- c2c0812[[i]] %>%
    filter(!row_number() %in% c(1, 2))
}

# Bind datasets together by year
c2c1519 <- bind_rows(c2c1519)
c2c1216 <- bind_rows(c2c1216)
c2c0812 <- bind_rows(c2c0812)
```


```{r, warning = FALSE, message = FALSE}
# Combine county migration data with controls and election data
df_2020 <- c2c1519 %>%
  filter(is.na(fips_b) == FALSE) %>%
  mutate(comb_fips_a = paste0(state_code_a, fips_a)) %>%
  group_by(county_name_a, comb_fips_a) %>%
  summarize(flow_ba = sum(as.numeric(flow_ba))) %>%
  arrange(comb_fips_a) %>%
  mutate(fips = as.numeric(comb_fips_a))

countypres2020 <- read_csv("countypres_2000-2020.csv") %>%
  filter(year == 2020) %>%
  filter(party == "DEMOCRAT" | party == "REPUBLICAN") %>%
  group_by(county_fips) %>%
  summarize(total_2p_votes = sum(candidatevotes)) %>%
  mutate(fips = as.numeric(county_fips)) %>%
  select(fips, total_2p_votes)

countypres2020_2p <- read_csv("countypres_2000-2020.csv") %>%
  filter(year == 2020) %>%
  filter(party == "DEMOCRAT") %>%
  mutate(fips = as.numeric(county_fips)) %>%
  group_by(fips) %>%
  summarize(candidatevotes = sum(candidatevotes)) %>%
  left_join(countypres2020, by = "fips") %>%
  mutate(dem_percent = candidatevotes/total_2p_votes)

countypres2020_2p_rep <- read_csv("countypres_2000-2020.csv") %>%
  filter(year == 2020) %>%
  filter(party == "REPUBLICAN") %>%
  mutate(fips = as.numeric(county_fips)) %>%
  group_by(fips) %>%
  summarize(candidatevotes = sum(candidatevotes)) %>%
  left_join(countypres2020, by = "fips") %>%
  mutate(rep_percent = candidatevotes/total_2p_votes) %>%
  select(fips, rep_percent)
  

df_2020_election <- left_join(df_2020, countypres2020_2p, by = "fips")

df_2020_combined <- left_join(df_2020_election, controls_2019, by = "fips") %>%
  mutate(avg_immi_rate = flow_ba/total_pop)
```



```{r}
df <- df_2020_combined %>%
  filter(Geo_STATE != "02" && Geo_STATE != "72")
```

```{r}
write.csv(df, "migration_df.csv")
```


```{r}
countypres2020_2p %>%
  filter(fips > 2000) %>%
  filter(fips < 3000) %>%
  group_by() %>%
  summarize(candidate_votes = sum(candidatevotes), total_2p_votes = sum(total_2p_votes))
```



# Immigration

```{r, warning = FALSE, message = FALSE}
# Basic linear regression
stargazer(lm(dem_percent ~ avg_immi_rate, data = df_2020_combined, weight = total_pop),
          type = "text")
```

```{r, warning = FALSE, message = FALSE}
# Visualization of relationship between immigration rate and dem percent
ggplot(df_2020_combined, aes(x = avg_immi_rate, y = dem_percent, size = total_pop)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", aes(weight = total_pop), show.legend = FALSE) +
  labs(title = "Fall in Democratic Voting with Increase in Rate of Immigration",
       x = "Immigration Rate",
       y = "Two-Party Democratic Vote Share",
       subtitle = "2020 Presidential Election",
       caption = "Weighted by County Population")
```

Check for missing value in the economic data
Check Alaska!!
What are the patterns we expect to see if this is true, does the data look like what it would if this is happening
Have a hypothesis that this should impact behavior, do you see a pattern that isn't explained by other things
Matching
Viriginia reports cities and counties around them differently and doesn't line up - drop ones that don't align cleanly (fips codes matched things are weird, census has the full county but the data is reported split between the city and everything else)
Know things about the places that people are coming from, -1, 0, 1 net partisanship of the people moving into an area, sorting, impact on people there

```{r, warning = FALSE, message = FALSE}
# Regression with controls
stargazer(lm(dem_percent ~ avg_immi_rate + percent_white + percent_hispanic +
               percent_bachelors + avg_wage + poverty + unemp + med_household_income, data = df_2020_combined,
             weight = total_pop),
          lm(dem_percent ~ avg_immi_rate + percent_white + percent_hispanic +
               percent_bachelors, data = df_2020_combined,
             weight = total_pop),
          lm(dem_percent ~ avg_immi_rate + avg_wage + poverty + unemp + med_household_income, data = df_2020_combined,
             weight = total_pop),
          type = "text")
```


```{r, warning = FALSE, message = FALSE}
# Regression standardized
df_2020_combined$dem_percent_scaled <- scale(df_2020_combined$dem_percent)[, 1]
df_2020_combined$avg_immi_rate_scaled <- scale(df_2020_combined$avg_immi_rate)[, 1]
df_2020_combined$percent_white_scaled <- scale(df_2020_combined$percent_white)[, 1]
df_2020_combined$percent_bachelors_scaled <- scale(df_2020_combined$percent_bachelors)[, 1]
df_2020_combined$med_household_income_scaled <- scale(df_2020_combined$med_household_income)[, 1]
df_2020_combined$percent_hispanic_scaled <- scale(df_2020_combined$percent_hispanic)[, 1]
df_2020_combined$avg_wage_scaled <- scale(df_2020_combined$avg_wage)[, 1]
df_2020_combined$poverty_scaled <- scale(df_2020_combined$poverty)[, 1]
df_2020_combined$unemp_scaled <- scale(df_2020_combined$unemp)[, 1]

stargazer(lm(dem_percent_scaled ~ avg_immi_rate_scaled + percent_white_scaled + percent_bachelors_scaled + 
               med_household_income_scaled + percent_hispanic_scaled +
               avg_wage_scaled + poverty_scaled + unemp_scaled, data = df_2020_combined,
          weight = total_pop), type = "text")
```

# Emigration

```{r}
df_2020_emi <- c2c1519 %>%
  filter(is.na(fips_b) == FALSE) %>%
  mutate(comb_fips_a = paste0(state_code_a, fips_a)) %>%
  group_by(county_name_a, comb_fips_a) %>%
  summarize(flow_ab = sum(as.numeric(flow_ab))) %>%
  arrange(comb_fips_a) %>%
  mutate(fips = as.numeric(comb_fips_a))

df_2020_election_emi <- left_join(df_2020_emi, countypres2020_2p, by = "fips")

df_2020_combined_emi <- left_join(df_2020_election_emi, controls_2019, by = "fips") %>%
  mutate(avg_emi_rate = flow_ab/total_pop)
```

```{r}
df_2020_combined_emi %>%
  filter(Geo_STATE == "02") %>%
  group_by() %>%
  summarize(flow_ab = sum(flow_ab))
```


```{r}
df_emi <- df_2020_combined_emi %>%
  filter(Geo_STATE != "02") %>%
  filter(Geo_STATE != "72")
write.csv(df_emi, "migration_df2.csv")
```


```{r, warning = FALSE, message = FALSE}
# Basic regression
stargazer(lm(dem_percent ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop), type = "text")
```

```{r}
ggplot(df_2020_combined_emi, aes(x = avg_emi_rate, y = dem_percent, size = total_pop)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", aes(weight = total_pop), show.legend = FALSE) +
  labs(title = "Fall in Democratic Voting with Increase in Rate of Emigration",
       x = "Emigration Rate",
       y = "Two-Party Democratic Vote Share",
       subtitle = "2020 Presidential Election",
       caption = "Weighted by County Population")
```

```{r, warning = FALSE, message = FALSE}
stargazer(lm(dem_percent ~ avg_emi_rate + percent_white + percent_hispanic +
               percent_bachelors + avg_wage + poverty + unemp + med_household_income, data = df_2020_combined_emi,
             weight = total_pop), 
          lm(dem_percent ~ avg_emi_rate + percent_white + percent_hispanic +
               percent_bachelors, data = df_2020_combined_emi,
             weight = total_pop),
          lm(dem_percent ~ avg_emi_rate + avg_wage + poverty + unemp + med_household_income, data = df_2020_combined_emi,
             weight = total_pop),
          type = "text")
```

# Maps

```{r}
plot_usmap(data = df_2020_combined, values = "avg_immi_rate", size = .1) +
  scale_fill_continuous(low = "white", "high" = "black") +
  labs(title = "Internal Immigration Rates from 2015-2019")
```

```{r}
plot_usmap(data = df_2020_combined_emi, values = "avg_emi_rate", size = .1) +
  scale_fill_continuous(low = "white", "high" = "black") +
  labs(title = "Internal Emigration Rates from 2015-2019")
```

# Regressing on Controls

```{r}
stargazer(lm(percent_white ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop),
          lm(percent_hispanic ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop),
          lm(percent_bachelors ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop),
          type = "text")
```

```{r}
stargazer(lm(percent_white ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop),
          lm(percent_hispanic ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop),
          lm(percent_bachelors ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop),
          type = "text")
```

```{r}
stargazer(lm(avg_wage ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop),
          lm(poverty ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop),
          lm(unemp ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop),
          lm(med_household_income ~ avg_emi_rate, data = df_2020_combined_emi, weight = total_pop),
          type = "text")
```


Thoughts - add cost of living interacting with avg_wage?

```{r}
stargazer(lm(avg_immi_rate ~ avg_wage, data = df_2020_combined, weight = total_pop),
          lm(avg_immi_rate ~ poverty, data = df_2020_combined, weight = total_pop),
          lm(avg_immi_rate ~ unemp, data = df_2020_combined, weight = total_pop),
          lm(avg_immi_rate ~ med_household_income, data = df_2020_combined, weight = total_pop),
          type = "text")
```


```{r}
ggplot(df_2020_combined, aes(x = avg_immi_rate, y = percent_white, size = total_pop)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", aes(weight = total_pop), show.legend = FALSE) +
  labs(title = "Immigration Rate is Higher in Counties with Higher White Proportion",
       x = "Immigration Rate",
       y = "Proportion White",
       subtitle = "2019",
       caption = "Weighted by County Population")
```

```{r}
ggplot(df_2020_combined, aes(x = avg_immi_rate, y = percent_hispanic, size = total_pop)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", aes(weight = total_pop), show.legend = FALSE) +
  labs(title = "Immigration Rate is Lower in Counties with Higher Hispanic Proportion",
       x = "Immigration Rate",
       y = "Proportion Hispanic",
       subtitle = "2019",
       caption = "Weighted by County Population")
```

```{r}
ggplot(df_2020_combined_emi, aes(x = avg_emi_rate, y = percent_white, size = total_pop)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", aes(weight = total_pop), show.legend = FALSE) +
  labs(title = "Emigration Rate is Higher in Counties with Higher White Proportion",
       x = "Emigration Rate",
       y = "Proportion White",
       subtitle = "2019",
       caption = "Weighted by County Population")
```

```{r}
ggplot(df_2020_combined_emi, aes(x = avg_emi_rate, y = percent_hispanic, size = total_pop)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", aes(weight = total_pop), show.legend = FALSE) +
  labs(title = "Emigration Rate is Lower in Counties with Higher Hispanic Proportion",
       x = "Emigration Rate",
       y = "Proportion Hispanic",
       subtitle = "2019",
       caption = "Weighted by County Population")
```

```{r}
ggplot(df_2020_combined, aes(x = avg_immi_rate, y = percent_bachelors, size = total_pop)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", aes(weight = total_pop), show.legend = FALSE) +
  labs(title = "Immigration Rate is Higher in Counties with Higher Bachelor's Proportion",
       x = "Immigration Rate",
       y = "Percent Bachelor's Degree",
       subtitle = "2019",
       caption = "Weighted by County Population")
```

```{r}
ggplot(df_2020_combined_emi, aes(x = avg_emi_rate, y = percent_bachelors, size = total_pop)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", aes(weight = total_pop), show.legend = FALSE) +
  labs(title = "Emigration Rate is Higher in Counties with Higher Bachelor's Proportion",
       x = "Emigration Rate",
       y = "Percent Bachelor's Degree",
       subtitle = "2019",
       caption = "Weighted by County Population")
```


